A Novel Model-Free Learning Approach for Dynamic Target Tracking Problem and its Validation using Virtual Robot Experimentation Platform

### 2020-06-29 paperMIC2020/paperIEEE-MIC2020-V1.tex

\author{%~
\IEEEauthorblockN{Amr Elhussein and Md Suruz Miah}
\IEEEauthorblockA{\textit{Electrical \& Computer Eng.} \\
\textit{Bradley University}, Peoria, Illinois, USA \\
E-mails: aelhussein@mail.bradley.edu and smiah@bradley.edu}
% \and
% \IEEEauthorblockN{Fazel Keshtkar}
% \IEEEauthorblockA{\textit{Div. of Computer Science, Math. and Science} \\
% \textit{St John's University}, Queens, NY, USA \\
% keshtkaf@stjohns.edu}
% \and
% \IEEEauthorblockN{Mohammed Abouheaf}
% \IEEEauthorblockA{\textit{School of EECS} \\
% \textit{Univ. of Ottawa}, Canada\\
% mabouhea@uottawa.ca}
%
}

@@

% Note that, the policy iteration computational setup rearranges the temporal difference expression~\eqref{eq:valueFunctionEstimated} such that %
 %
 \begin{equation}
 \label{eq:const}
 \mathbf{z}_k^T\, \mathbf{P}\, \mathbf{z}_k - \mathbf{z}_{k+1}^T\, \mathbf{P}\, \mathbf{z}_{k+1} = \mathbf{z}_k^T\, \bar{\mathbf{P}}\, \mathbf{z}_k. 
 \end{equation}
 %
 This equation is utilized repeatedly in order to evaluate a certain policy during at least $\eta\ge \bar n, \bar n= (3+2)(3+2+1)/2$ evaluation steps (i.e., the lowest evaluation interval spans $k$ to $k+\bar n$ calculation samples) in order to update the critic weights vector  $\boldsymbol{\omega}=\mathrm{vec}(\mathbf{P}),$ which consists of connection weights between the neurons of the hidden layer and the output layer of the critic neural network shown in Fig.~\eqref{fig:nnCritic}. The operator $\mathrm{vec}(\mathbf{P})$ forms the columns of the $\mathbf{P}$ matrix into a column vector $\mathbf{\omega}$ of dimension $\bar{n}=15$ since the matrix $\mathbf{P}$ is a symmetric matrix. The left hand side of~\eqref{eq:const} is expressed using the following critic approximation form %
 %
 $$\hat{V}(\mathbf{z}_k)-\hat{V}(\mathbf{z}_{k+1})=\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1}),$$
 %
 where $\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1})=\bm{\rho}(\mathbf{z}_k)-\bm{\rho}(\mathbf{z}_{k+1}) \in \mathbb{R}^{15 \times 1}, \, \bm{\rho}(\mathbf{z}_k)=\left(\mathbf{z}^q_k\otimes\mathbf{z}^h_k\right)$ $(q=1,\dots, 5, \,\, h=q,\dots,5),$ and $\boldsymbol{\omega}^T = [0.5 \, P^{11},P^{12},P^{13},P^{14}, P^{15}, \, 0.5 \, P^{22}, P^{23},P^{24},P^{25},$ $\,0.5 \, P^{33}, P^{34},P^{35},\,0.5 \, P^{44},P^{45}, \,0.5 \, P^{55}]^T \in \mathbb{R}^{1\times 15}$ ($P^{ij}$ is the $ij^{th}$ entry of matrix $\mathbf{P}$). %
 %
 The critic weights $\boldsymbol{\omega}$ are updated using a gradient descent approach, where the tuning error $\varepsilon_k$ at each computational instance $k$ follows $\varepsilon_k =\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1})-{v}_k,$ where $v_\kappa = \frac{1}{2}\mathbf{z}_{k}^T \, \bar{\mathbf{P}} \, \mathbf{z}_{k}$. As detailed earlier, it is required to perform at least $\eta \ge \bar n$ evaluation steps before updating the critic weights $\boldsymbol{\omega}$ (i.e., finding the new improved policy). Hence, it is required to minimize the sum of square errors such that %
 %
 \begin{multline*}
   \delta_c =\sum_{\kappa=0}^{\eta-1}\frac{1}{2}(\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k+\kappa,k+\kappa+1})-{v}_{k+\kappa})^2 = \frac{1}{2}\| \mathbf{v} - \bm{\Lambda}\bm{\omega}\|^2\\
   =\frac{1}{2}\left(\mathbf{v} - \bm{\Lambda}\bm{\omega}\right)^T \left(\mathbf{v} - \bm{\Lambda}\bm{\omega}\right), 
 \end{multline*}
 where $\bm{\Lambda} = [{\bf o}_0,{\bf o}_1,\ldots,{\bf o}_{\eta-1}]^T \in \mathbb{R}^{\eta\times 15}$ with ${\bf o}_\kappa = \tilde{\bm{\rho}}^T({\bf z}_{k+\kappa, k+\kappa+1}) \in \mathbb{R}^{1\times 15}$ and ${\bf v} =[v_0,v_1,\ldots,v_{\eta-1}]^T \in \mathbb{R}^{\eta}$ with $v_\kappa = \frac{1}{2}\mathbf{z}_{k+\kappa}^T \, \bar{\mathbf{P}} \, \mathbf{z}_{k+\kappa}$ for $\kappa = 0,1,\ldots, \eta-1$. %
 Therefore, the update law of the critic weights using the gradient decent approach for at least $\bar n$ samples is given by %
 \begin{multline}
   \bm{\omega}^{[r+1]} = \bm{\omega}^{[r]} - \ell_c\frac{\partial\delta_c}{\partial \bm{\omega}} = \bm{\omega}^{[r]} - \ell_c\left(-\bm{\Lambda}^T\mathbf{v} + \bm{\Lambda}^T \bm{\Lambda}\bm{\omega}^{[r]}\right)\\ 
 =\bm{\omega}^{[r]} - \ell_c \bm{\Lambda}^T\left(\bm{\Lambda}\bm{\omega}^{[r]}-\mathbf{v}\right), 
 \label{eq:criticWeights}
 \end{multline}
 %
 where $0<\ell_c<1$ is a critic learning rate and $r$ is the  update index of the critic weights.

% The newly computed critic weights $\boldsymbol{\omega}$ are used to reconstruct the matrix ${\bf P}$ (i.e., updating the solving value function and hence calculating the associated policy) such that 
 %
 \begin{center}
 ${\bf P}=\begin{bmatrix} 
 2\,\omega^{1}    & \omega^{2}      & \omega^{3}       & \omega^{4}       & \omega^{5} \\ 
 \omega^{2}       &2\, \omega^{6}   & \omega^{7}       & \omega^{8}       & \omega^{9} \\
 \omega^{3}       & \omega^{7}      &2\, \omega^{10}   & \omega^{11}      & \omega^{12} \\
 \omega^{4}       & \omega^{8}      & \omega^{11}      & 2\,\omega^{13}   & \omega^{14} \\
 \omega^{5}       & \omega^{9}      & \omega^{12}      & \omega^{14}      & 2\,\omega^{15}   
 \end{bmatrix}\in \mathbb{R}^{5\times 5},$
 \end{center}
 %
 where $\omega^{i}$ is the $i^{th}$ entry of the weight vector $\bm{\omega}.$ The complete policy iteration solution process for the leader-follower problem is detailed out in Algorithm \ref{alg:ModelFreeTracking}. %
 %
 \begin{algorithm2e}
     \caption{\label{alg:ModelFreeTracking} Model-free reinforcement learning using the policy iteration solution.}
     \DontPrintSemicolon
     \KwIn{Sampling-time $T_s,$ $\mathbf{Q},~\text{and}~\mathbf{R}$ }
     \KwOut{Error trajectory $\mathbf{e}_k,$ for $k=0,1,\ldots$}
     % \KwData{Map representation using occupancy grid technique}
     % \KwResult{output******************************}
     \Begin{
         $k=0, r = 0$ \tcc*[h]{Discrete time and policy indices}\;
         $\eta = (n+m)(n+m+1)/2$\;
         Initialize $\mathbf{P}^{[0]}$ \tcc*[h]{Positive definite}\;
                 Set offset distance $d$\;
                 Given approximate initial poses of leader and follower, compute $\mathbf{e}_0$ using error model~\eqref{eq:stateError}\;
                 Compute follower's input $\mathbf{u}_0^{[0]}$ using policy~\eqref{eq:modelFreePolicy}\;
         \Repeat(\tcc*[h]{Main timing loop}){Tracking errors are zero}
         {
             Find $\mathbf{e}_{k+1}$ using~\eqref{eq:stateError}\;
             Compute policy $\mathbf{u}_{k+1}^{[r]}$ using~\eqref{eq:modelFreePolicy}\;
             %Evaluate and record Equation~\eqref{eq:modelFreeMatrixSolution}\;
             \If{[$(k+1)~\mathrm{modulo}~\eta]==0$ }
             {
                 $r\leftarrow r+1$\tcc*[h]{Evaluate policy}\;
                 Solve for the critic-weights $\bm{\omega}$ using~\eqref{eq:criticWeights}\;
                 Construct matrix $\mathbf{P}^{[r]}$ using vector $\bm{\omega}$\;
 %                \eIf{$\|\mathbf{P}^{[r]} - \mathbf{P}^{[r+1]}\|<\varepsilon$}{Set $\mathbf{u}_{k+1}^*\leftarrow\mathbf{u}_{k+1}^{[r]}$}{$k\leftarrow k+1$ }
             }
 %        {
 %                $k\leftarrow k+1$
 %            }
         $k\leftarrow k+1$
         }
     }
 \end{algorithm2e}

>>>>

\author{%~
\IEEEauthorblockN{Amr Elhussein and Md Suruz Miah}
\IEEEauthorblockA{\textit{Electrical \& Computer Eng.} \\
\textit{Bradley University}, Peoria, Illinois, USA \\
 aelhussein@mail.bradley.edu and smiah@bradley.edu}
% \and
% \IEEEauthorblockN{Fazel Keshtkar}
% \IEEEauthorblockA{\textit{Div. of Computer Science, Math. and Science} \\
% \textit{St John's University}, Queens, NY, USA \\
% keshtkaf@stjohns.edu}
% \and
% \IEEEauthorblockN{Mohammed Abouheaf}
% \IEEEauthorblockA{\textit{School of EECS} \\
% \textit{Univ. of Ottawa}, Canada\\
% mabouhea@uottawa.ca}
%
}

@@

% Note that, the policy iteration computational setup rearranges the temporal difference expression~\eqref{eq:valueFunctionEstimated} such that %
 %
 \begin{equation}
 \label{eq:const}
 \mathbf{z}_k^T\, \mathbf{P}\, \mathbf{z}_k - \mathbf{z}_{k+1}^T\, \mathbf{P}\, \mathbf{z}_{k+1} = \mathbf{z}_k^T\, \bar{\mathbf{P}}\, \mathbf{z}_k. 
 \end{equation}
 %
 This equation is utilized repeatedly in order to evaluate a certain policy during at least $\eta\ge \bar n, \bar n= (2+2)(2+2+1)/2$ evaluation steps (i.e., the lowest evaluation interval spans $k$ to $k+\bar n$ calculation samples) in order to update the critic weights vector  $\boldsymbol{\omega}=\mathrm{vec}(\mathbf{P}),$ which consists of connection weights between the neurons of the hidden layer and the output layer of the critic neural network shown in Fig.~\eqref{fig:nnCritic}. The operator $\mathrm{vec}(\mathbf{P})$ forms the columns of the $\mathbf{P}$ matrix into a column vector $\mathbf{\omega}$ of dimension $\bar{n}=15$ since the matrix $\mathbf{P}$ is a symmetric matrix. The left hand side of~\eqref{eq:const} is expressed using the following critic approximation form %
 %
 $$\hat{V}(\mathbf{z}_k)-\hat{V}(\mathbf{z}_{k+1})=\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1}),$$
 %
 where $\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1})=\bm{\rho}(\mathbf{z}_k)-\bm{\rho}(\mathbf{z}_{k+1}) \in \mathbb{R}^{10 \times 1}, \, \bm{\rho}(\mathbf{z}_k)=\left(\mathbf{z}^q_k\otimes\mathbf{z}^h_k\right)$ $(q=1,\dots, 5, \,\, h=q,\dots,5),$ and $\boldsymbol{\omega}^T = [0.5 \, P^{11},P^{12},P^{13},P^{14},  \, 0.5 \, P^{22}, P^{23},P^{24},$ $\,0.5 \, P^{33}, P^{34},\,0.5 \ P^{44}, ]^T \in \mathbb{R}^{1\times 15}$ ($P^{ij}$ is the $ij^{th}$ entry of matrix $\mathbf{P}$). %
 %
 The critic weights $\boldsymbol{\omega}$ are updated using a gradient descent approach, where the tuning error $\varepsilon_k$ at each computational instance $k$ follows $\varepsilon_k =\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1})-{v}_k,$ where $v_\kappa = \frac{1}{2}\mathbf{z}_{k}^T \, \bar{\mathbf{P}} \, \mathbf{z}_{k}$. As detailed earlier, it is required to perform at least $\eta \ge \bar n$ evaluation steps before updating the critic weights $\boldsymbol{\omega}$ (i.e., finding the new improved policy). Hence, it is required to minimize the sum of square errors such that %
 %
 \begin{multline*}
   \delta_c =\sum_{\kappa=0}^{\eta-1}\frac{1}{2}(\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k+\kappa,k+\kappa+1})-{v}_{k+\kappa})^2 = \frac{1}{2}\| \mathbf{v} - \bm{\Lambda}\bm{\omega}\|^2\\
   =\frac{1}{2}\left(\mathbf{v} - \bm{\Lambda}\bm{\omega}\right)^T \left(\mathbf{v} - \bm{\Lambda}\bm{\omega}\right), 
 \end{multline*}
 where $\bm{\Lambda} = [{\bf o}_0,{\bf o}_1,\ldots,{\bf o}_{\eta-1}]^T \in \mathbb{R}^{\eta\times 15}$ with ${\bf o}_\kappa = \tilde{\bm{\rho}}^T({\bf z}_{k+\kappa, k+\kappa+1}) \in \mathbb{R}^{1\times 15}$ and ${\bf v} =[v_0,v_1,\ldots,v_{\eta-1}]^T \in \mathbb{R}^{\eta}$ with $v_\kappa = \frac{1}{2}\mathbf{z}_{k+\kappa}^T \, \bar{\mathbf{P}} \, \mathbf{z}_{k+\kappa}$ for $\kappa = 0,1,\ldots, \eta-1$. %
 Therefore, the update law of the critic weights using the gradient decent approach for at least $\bar n$ samples is given by %
 \begin{multline}
   \bm{\omega}^{[r+1]} = \bm{\omega}^{[r]} - \ell_c\frac{\partial\delta_c}{\partial \bm{\omega}} = \bm{\omega}^{[r]} - \ell_c\left(-\bm{\Lambda}^T\mathbf{v} + \bm{\Lambda}^T \bm{\Lambda}\bm{\omega}^{[r]}\right)\\ 
 =\bm{\omega}^{[r]} - \ell_c \bm{\Lambda}^T\left(\bm{\Lambda}\bm{\omega}^{[r]}-\mathbf{v}\right), 
 \label{eq:criticWeights}
 \end{multline}
 %
 where $0<\ell_c<1$ is a critic learning rate and $r$ is the  update index of the critic weights.

% The newly computed critic weights $\boldsymbol{\omega}$ are used to reconstruct the matrix ${\bf P}$ (i.e., updating the solving value function and hence calculating the associated policy) such that 
 %
 \begin{center}
 ${\bf P}=\begin{bmatrix} 
 2\,\omega^{1}    & \omega^{2}      & \omega^{3}       & \omega^{4}        \\ 
 \omega^{2}       &2\, \omega^{5}   & \omega^{6}       & \omega^{7}        \\
 \omega^{3}       & \omega^{6}      &2\, \omega^{8}   & \omega^{9}   \\   
 \omega^{4}       & \omega^{7}      & \omega^{9}      & 2\,\omega^{10}      
 \end{bmatrix}\in \mathbb{R}^{4\times 4},$
 \end{center}
 %
 where $\omega^{i}$ is the $i^{th}$ entry of the weight vector $\bm{\omega}.$ 
 \todo[inline]{merge the actor weights discussion with the rest}
 The critic wegihts are then used to update the actor weights which maps the state error to the desired ploicy (control actions) at every time step $t$ with the following equation:
 \begin{align}
 \label{eq:actorWeightsPolicy}    
 \mathbf{u}_k  =  \mathbf{\omega_a^{T}} \mathbf{e}_k
 \end{align}   
 Applying a similar gradient descent approach as with the critic weights the actor weights are updated as follows:
\begin{align}
   \bm{\omega_a}^{[r+1]} = \bm{\omega}^{[r]} - \ell_a(\mathbf{u}_k-\mathbf{u}_k^*)\mathbf{e}_k, 
 \label{eq:actorWeights}
 \end{align}
 where $\ell$ is the actor learning rate. 
 \\
 The complete policy iteration solution process for the leader-follower problem is detailed out in Algorithm \ref{alg:ModelFreeTracking}. %
 %
 \begin{algorithm2e}{
     \caption{\label{alg:ModelFreeTracking} Model-free actor-critic reinforcement learning using the policy iteration solution.}
     \DontPrintSemicolon
     \KwIn{Sampling-time $T_s,$ $\mathbf{Q},~\text{and}~\mathbf{R}$ }
     \KwOut{Error trajectory $\mathbf{e}_k,$ for $k=0,1,\ldots$}
     % \KwData{Map representation using occupancy grid technique}
     % \KwResult{output******************************}
     \Begin{
         $k=0, r = 0$ \tcc*[h]{Discrete time and policy indices}\;
         $\eta = (n+m)(n+m+1)/2$\;
         Initialize $\mathbf{P}^{[0]}$ \tcc*[h]{Positive definite}\;
                 Set offset distance $d$\;
                 Given approximate initial poses of leader and follower, compute $\mathbf{e}_0$ using error model~\eqref{eq:stateError}\;
                 Compute follower's input $\mathbf{u}_0^{[0]}$ using policy~\eqref{eq:actorWeightsPolicy}\;
         \Repeat(\tcc*[h]{Main timing loop}){Tracking errors are zero}
         {
             Find $\mathbf{e}_{k+1}$ using~\eqref{eq:stateError}\;
             Compute policy $\mathbf{u}_{k+1}^{[r]}$ using~\eqref{eq:actorWeightsPolicy}\;
             %Evaluate and record Equation~\eqref{eq:modelFreeMatrixSolution}\;
             \If{[$(k+1)~\mathrm{modulo}~\eta]==0$ }
             {
                 $r\leftarrow r+1$\tcc*[h]{Evaluate policy}using~\eqref{eq:modelFreePolicy}\;
                 Solve for the critic-weights $\bm{\omega_c}$ using~\eqref{eq:criticWeights}eights}\;
                 Solve for the actor-weights $\bm{\omega_a}$ using ~\eqref{eq:actorWeights}\;\;
                 Construct matrix $\mathbf{P}^{[r]}$ using vector $\bm{\omega_c}$\;
 %                \eIf{$\|\mathbf{P}^{[r]} - \mathbf{P}^{[r+1]}\|<\varepsilon$}{Set $\mathbf{u}_{k+1}^*\leftarrow\mathbf{u}_{k+1}^{[r]}$}{$k\leftarrow k+1$ }
             }
 %        {
 %                $k\leftarrow k+1$
 %            }
         $k\leftarrow k+1$
         }
     }
 \end{algorithm2e}
